---
title: "Airline, Airport choice - EDA"
author: "Huy Le"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

Commented header  
=========================================

```{r echo = TRUE}
# Course: 
# Title: 
# Purpose: 
# Date: 
# Author: Huy LE
```


Clear environment of variables and packages  
=========================================

```{r}

rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

Load packages that include different functions  
=========================================


```{r }
# Load libraries
library(tidyverse)
library(GGally)
library(gridExtra)
library("readxl")
library(janitor)


```

Load data and begin new EDA  
=========================================


```{r}
# Load data 
df <- read_excel("./raw_data/airport_choice_survey_EN_ver2.0_Capstone.xlsx")
df1<-df[-c(1,13,15,16,27)]
# Examine data structure 
str(df1)
```

```{r}
df <- df %>% mutate(Airport = ifelse(Airport==1,0,1))
dat <- df %>% select(-ID) %>%
        mutate(Airport = factor(ifelse(Airport==1, 0 ,1), levels= c(0,1)),
               Airline = factor(Airline, levels= c(1,2,3,4)),
               Gender = factor(Gender, levels= c(1,2)),
               Nationality = factor(Nationality, levels= c(1,2,3,4,5)),
               TripPurpose = factor(TripPurpose, levels= c(1,2,3,4)),
               ProvinceResidence = factor(ProvinceResidence, levels= c(1,2,3,4,5,6,7,8)),
               GroupTravel = factor(GroupTravel, levels= c(1,2)),
               FrequentFlightDestination = factor(FrequentFlightDestination, levels= c(1,2,3,4,5,6,7)),
               Destination = factor(Destination, levels= c(1,2,3,4)),
               DepartureTime = factor(DepartureTime, levels= c(1,2,3,4)),
               SeatClass = factor(SeatClass, levels= c(1,2,3)),
               ModeTransport = factor(ModeTransport, levels= c(1,2,3,4,5,6,7,8,9,10,11)),
               Occupation = factor(Occupation, levels= c(1,2,3,4,5,6,7,8,9,10,11,12)),
               Income = factor(Income, levels= c(1,2,3,4,5,6,7)),
               MileageAirline = factor(MileageAirline, levels= c(1,2,3,4,5)),
               DepartureHr = as.integer(DepartureHr)) %>% as.data.frame()

summary(dat)
typeof(dat)
#check na value
sapply(df, function(x) sum(is.na(x)))

sapply(df, function(x) sum(is.infinite(x)))
```

               ID                   Airport                   Airline                       Age                    Gender 
                        0                         0                        10                         1                         3 
              Nationality               TripPurpose              TripDuration           FlyingCompanion         ProvinceResidence 
                        0                         0                         0                         0                         0 
              GroupTravel           NoTripsLastYear FrequentFlightDestination               Destination                  FlightNo 
                        0                         0                        32                         5                       142 
              DepartureHr               DepartureMn             DepartureTime                 SeatClass                   Airfare 
                       33                       120                         0                         4                       155 
              NoTransport             ModeTransport                AccessCost                AccessTime                Occupation 
                        0                         0                       197                        97                         0 
                   Income            MileageAirline                   Mileage 
                      132                       251                       398 

1. Univariate non-graphical -- Categorical
=========================================
##Step 1

```{r}
count_hist<- function(plot){
  plot + geom_histogram(bins = 52)+
    theme_bw() + 
    theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank()) + 
    labs(y ="Count")
}

count_hist(ggplot(df, aes(Mileage)))

count_bargraph <- function(plot) {
 plot + 
    geom_bar(position = "dodge") +
    theme_bw() + 
    theme(panel.border = element_blank(), 
                       panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank()) + 
    labs(y ="Count")
}
count_bargraph(ggplot(df, aes(Nationality)))
```

##Step 2

```{r}
ggpairs(df[,c("Age", "TripDuration", "NoTripsLastYear", "DepartureHr", "Mileage","Airfare", "NoTransport","AccessCost","AccessTime")])

ggpairs(df)
```

##Step 3


```{r}
df1 <- df %>% select(-FlightNo)

cor(df[,c("Age", "TripDuration", "NoTripsLastYear", "DepartureHr", "Mileage","Airfare", "NoTransport","AccessCost","AccessTime")], 
      use = "complete.obs")
```

```{r}
#Function for table
count_crosstable <- function(x, y, name){
   x = enquo(x)
   y = enquo(y)
   dat %>%
      tabyl(!!x, !!y) %>%
      adorn_totals(where = c('row','col'))
      #adorn_pct_formatting(digits = 0 ) %>%
}

count_crosstable(Airport, Airline,"airport vs Airline")
count_crosstable(Airport, Gender,"airport vs Gender")
count_crosstable(Airport, Nationality,"airport vs Nationality")
count_crosstable(Airport, TripPurpose,"airport vs TripPurpose")
count_crosstable(Airport, ProvinceResidence,"airport vs Tripduation")
count_crosstable(Airport, Income,"airport vs Tripduation")
```
Airport   1   2  3   4 NA_ Total
       1  76  54 57  44   8   239
       2  77  53 24  93   2   249
   Total 153 107 81 137  10   488 
   
   
gender
Airport   1   2 NA_ Total
       1 115 122   2   239
       2 114 134   1   249
   Total 229 256   3   488
  
  Nationality
 Airport   1  2  3  4  5 Total
       1 205  7  1 19  7   239
       2 173 26 41  2  7   249
   Total 378 33 42 21 14   488

ProvinceResidence
Airport   1  2   3  4  5  6 7  8 Total
       1  99 11  68  8 26  5 2 20   239
       2  87 11  60  6  5  5 4 71   249
   Total 186 22 128 14 31 10 6 91   488

Income
Airport  1  2  3  4  5  6  7 NA_ Total
       0 10 33 39 22 11  7  6 111   239
       1 41 60 46 39 22 12  8  21   249
   Total 51 93 85 61 33 19 14 132   488
```{r}
dat %>% ggplot(aes(x = Airline, fill =Airport)) +
        geom_bar(stat = "count")
```
```{r}
dat %>% ggplot(aes(x = Airline,y= Airfare)) +
        geom_boxplot()
```

```{r}
dat %>% ggplot(aes(x = Airport, y = AccessTime)) +
        geom_boxplot()
dat %>% ggplot(aes(x = Airport, y = AccessCost)) +
        geom_boxplot()
dat %>% ggplot(aes(x = Airport, y = Age)) +
        geom_boxplot()

dat %>% ggplot(aes(x = Income, fill = Airport)) +
        geom_bar(stat = "count")

dat["AccessCost"]
```

- Most of peopel with various income choose airport 1
- People who do not reveal their income chose airport 0

```{r}
dat %>% ggplot(aes(x = Airfare, y = AccessCost, colour = Airport)) +
        geom_point(alpha =0.5)
```

```{r}
dat %>% ggplot(aes(x = Mileage, y = Airfare, colour = Airport)) +
        geom_point(alpha =0.5)
```

```{r}
airportlist <- c("Airline","Gender","Nationality","TripPurpose","ProvinceResidence","GroupTravel","FrequentFlightDestination","Destination","DepartureTime","SeatClass","ModeTransport","Occupation","Income","MileageAirline")

airline_list <- c("Airport","Gender","Nationality","TripPurpose","ProvinceResidence","GroupTravel","FrequentFlightDestination","Destination","DepartureTime","SeatClass","ModeTransport","Occupation","Income","MileageAirline")

numlist <- c("Age", "TripDuration", "NoTripsLastYear", "DepartureHr", "Mileage","Airfare", "NoTransport","AccessCost","AccessTime")

bar_chart <- function(listx, yaxis){
  #y = enquo(yaxis)
  for(i in 1:length(listx)){
    #x <- sym(listx[i])
    p<- dat %>% ggplot(aes_string(x = listx[i], fill= yaxis)) +
            geom_bar(stat = "count", position ="dodge") 
    print(p)
  }
}
bar_chart(airportlist, "Airport")


bar_chart(airline_list, "Airline")


dat %>% ggplot(aes(x = SeatClass, fill =Airport)) +
        geom_bar(stat = "count") +
        facet_grid(Airline ~ .)

dat %>% ggplot(aes(x = TripPurpose, fill =Airport)) +
        geom_bar(stat = "count") +
        facet_grid(Airline ~ .)
ModeTransport

dat %>% ggplot(aes(x = ModeTransport, fill =Airport)) +
        geom_bar(stat = "count", position = "dodge") 

dat %>% ggplot(aes(x = NoTransport, fill =Airport)) +
        geom_bar(stat = "count", position = "dodge") 

#DepartureTime
dat %>% ggplot(aes(x = DepartureTime, fill =Airport)) +
        geom_bar(stat = "count", position = "dodge") 

# departure time has strong effect on traveler choice


```

  
"Airport","Airline","Age","Gender","Nationality","TripPurpose","TripDuration","FlyingCompanion","ProvinceResidence","GroupTravel","NoTripsLastYear","FrequentFlightDestination","Destination","FlightNo","DepartureHr","DepartureMn","DepartureTime","SeatClass","Airfare","NoTransport","ModeTransport","AccessCost","AccessTime","Occupation","Income","MileageAirline","Mileage"
```{r}
library(imputeTS)
library(fastDummies)
df2 <- df

dummy_cols(df2, select_columns =- c("Age", "TripDuration", "NoTripsLastYear", "DepartureHr", "Mileage","Airfare", "NoTransport","AccessCost","AccessTime","FlightNo"), remove_first_dummy = TRUE)
typeof(df)

fit_linear <-glm(Airport ~ Airline + 
                   #Age +
                   #Gender +
                   #Destination + 
                   Income + 
                   #Mileage + 
                   AccessTime +
                   #ProvinceResidence+
                   AccessCost+
                   #DepartureTime +
                   Airfare 
                   #FrequentFlightDestination +
                   #ModeTransport
                 ,
                 data =dat, na.action="na.omit",
                 family = "binomial")

summary(fit_linear)

```
```{r}

```

